---
title: "LOTR Network"
author: "Christian"
date: "`r Sys.Date()`"
output: html_document
---

# One Network Visualisation to Rule Them 

## Table of contents

## Description

This Rmarkdown document will detail the process behind the visualisations made from this project. These include

 * Degree centrality for the Two Towers
 * Betweeness centrality for the Two Towers
 * Closeness centrality for the Two Towers
 * Degree centrality for the combined books
 * Betweeness centrality for the combined books
 * Closeness centrality for the combined books
 * Evolution of degree centrality accross the three books

```{r libraries, echo=FALSE, message=FALSE, warning=FALSE}

library(tidyverse)
library(reshape2)

```

## Degree centrality: The Two Towers

For this plot we'll import the data, take the first 10 characters by centrality, and visualise the data in a barplot

```{r dc_tt, echo=TRUE, fig.align='center', message=FALSE, warning=FALSE}

dc_tt <- read_csv('Centrality Measures TT/Centrality.csv') %>%
  rename(character = ...1) %>%
  arrange(desc(centrality)) %>%
  slice_head(n = 10)

tt_centrality <- ggplot(dc_tt, 
                        aes(x = reorder(character, desc(centrality)), 
                        y = centrality, 
                        fill = character)) +
  
  geom_bar(stat = "identity") +
  
  scale_y_continuous(
    limits = c(0,0.525),
    expand = c(0,0)
  ) +
  
  labs(x = NULL, y = "Centrality", title = "Character Importance (The Two Towers): Centrality", subtitle = "Top 10 most important characters in The Two Towers, as measured by degree centrality") +
  
  theme_classic() +
  
  theme(
    legend.position = "none", 
    panel.grid.major.y = element_line(color = "grey",
                                          size = 0.1,
                                          linetype = 1),
    panel.grid.minor.y = element_line(color = "grey",
                                          size = 0.1,
                                          linetype = 1)
  )
  
ggsave("Images/tt_centality.png", height = 4.47, width = 7.2)

tt_centrality
  
```

## Betweenness centrality: The Two Towers

This code will take the top 10 characters by betweenness centrality and visualise those using a bar chart.

```{r bc_tt, echo=TRUE, fig.align='center', message=FALSE, warning=FALSE}

bc_tt <- read_csv('Centrality Measures TT/Betweenness.csv') %>%
  rename(character = ...1) %>%
  arrange(desc(centrality)) %>%
  slice_head(n = 10)

tt_betweenness <- ggplot(bc_tt, 
                        aes(x = reorder(character, desc(centrality)), 
                        y = centrality, 
                        fill = character)) +
  
  geom_bar(stat = "identity") +
  
  scale_y_continuous(
    limits = c(0,0.2),
    expand = c(0,0)
  ) +
  
  labs(x = NULL, y = "Betweenness", title = "Character Importance (The Two Towers): Betweeness", subtitle = "Top 10 most important characters in The Two Towers, as measured by betweenness") +
  
  theme_classic() +
  
  theme(
    legend.position = "none", 
    panel.grid.major.y = element_line(color = "grey",
                                          size = 0.1,
                                          linetype = 1),
    panel.grid.minor.y = element_line(color = "grey",
                                          size = 0.1,
                                          linetype = 1)
  )
  
ggsave("Images/tt_betweenness.png", height = 4.47, width = 7.2)

tt_betweenness

```

## Closeness centrality: The Two Towers

This code will take the top 10 characters by closeness centrality and visualise those using a bar chart.

```{r cc_tt, echo=TRUE, fig.align='center', message=FALSE, warning=FALSE}

cc_tt <- read_csv('Centrality Measures TT/Closeness.csv') %>%
  rename(character = ...1) %>%
  arrange(desc(centrality)) %>%
  slice_head(n = 10)

tt_closeness <- ggplot(cc_tt, 
                        aes(x = reorder(character, desc(centrality)), 
                        y = centrality, 
                        fill = character)) +
  
  geom_bar(stat = "identity") +
  
  scale_y_continuous(
    limits = c(0,0.69),
    expand = c(0,0)
  ) +
  
  labs(x = NULL, y = "Betweenness", title = "Character Importance (The Two Towers): Closeness", subtitle = "Top 10 most important characters in The Two Towers, as measured by closeness") +
  
  theme_classic() +
  
  theme(
    legend.position = "none", 
    panel.grid.major.y = element_line(color = "grey",
                                          size = 0.1,
                                          linetype = 1),
    panel.grid.minor.y = element_line(color = "grey",
                                          size = 0.1,
                                          linetype = 1)
  )
  
ggsave("Images/tt_closeness.png", height = 4.47, width = 7.2)

tt_closeness

```


## Degree centrality: All Books

Next, we'll apply all of the centrality measures to the relationships from the combined books. Starting with the degree centrality measure.

```{r dc_all, echo=TRUE, fig.align='center', message=FALSE, warning=FALSE}

dc_all <- read_csv('Centrality Measures All/Centrality.csv') %>%
  rename(character = ...1) %>%
  arrange(desc(centrality)) %>%
  slice_head(n = 10)

all_centrality <- ggplot(dc_all, 
                        aes(x = reorder(character, desc(centrality)), 
                        y = centrality, 
                        fill = character)) +
  
  geom_bar(stat = "identity") +
  
  scale_y_continuous(
    limits = c(0,0.575),
    expand = c(0,0)
  ) +
  
  labs(
    x = NULL, 
    y = "Centrality", 
    title = "Character Importance (All Books): Centrality", 
    subtitle = "Top 10 most important characters in The Lord of the Rings, as measured by degree centrality"
  ) +
  
  theme_classic() +
  
  theme(
    legend.position = "none", 
    panel.grid.major.y = element_line(color = "grey",
                                          size = 0.1,
                                          linetype = 1),
    panel.grid.minor.y = element_line(color = "grey",
                                          size = 0.1,
                                          linetype = 1)
  )
  
ggsave("Images/all_centality.png", height = 4.47, width = 7.2)

all_centrality
  
```

## Betweenness centrality: All Books

This code will take the top 10 characters by betweenness centrality and visualise those using a bar chart.

```{r bc_all, echo=TRUE, fig.align='center', message=FALSE, warning=FALSE}

bc_all <- read_csv('Centrality Measures All/Betweenness.csv') %>%
  rename(character = ...1) %>%
  arrange(desc(centrality)) %>%
  slice_head(n = 10)

all_betweenness <- ggplot(bc_all, 
                        aes(x = reorder(character, desc(centrality)), 
                        y = centrality, 
                        fill = character)) +
  
  geom_bar(stat = "identity") +
  
  scale_y_continuous(
    limits = c(0,0.245),
    expand = c(0,0)
  ) +
  
  labs(x = NULL, 
       y = "Betweenness", 
       title = "Character Importance (All Books): Betweeness", 
       subtitle = "Top 10 most important characters in The Lord of the Rings, as measured by betweenness"
  ) +
  
  theme_classic() +
  
  theme(
    legend.position = "none", 
    panel.grid.major.y = element_line(color = "grey",
                                          size = 0.1,
                                          linetype = 1),
    panel.grid.minor.y = element_line(color = "grey",
                                          size = 0.1,
                                          linetype = 1)
  )
  
ggsave("Images/all_betweenness.png", height = 4.47, width = 7.2)

all_betweenness

```


## Closeness centrality: All Books

This code will take the top 10 characters by closeness centrality and visualise those using a bar chart.

```{r cc_all, echo=TRUE, fig.align='center', message=FALSE, warning=FALSE}

cc_all <- read_csv('Centrality Measures All/Closeness.csv') %>%
  rename(character = ...1) %>%
  arrange(desc(centrality)) %>%
  slice_head(n = 10)

all_closeness <- ggplot(cc_all, 
                        aes(x = reorder(character, desc(centrality)), 
                        y = centrality, 
                        fill = character)) +
  
  geom_bar(stat = "identity") +
  
  scale_y_continuous(
    limits = c(0,0.69),
    expand = c(0,0)
  ) +
  
  labs(x = NULL, y = "Betweenness", title = "Character Importance (All Books): Closeness", subtitle = "Top 10 most important characters in The Lord of the Rings, as measured by closeness") +
  
  theme_classic() +
  
  theme(
    legend.position = "none", 
    panel.grid.major.y = element_line(color = "grey",
                                          size = 0.1,
                                          linetype = 1),
    panel.grid.minor.y = element_line(color = "grey",
                                          size = 0.1,
                                          linetype = 1)
  )
  
ggsave("Images/all_closeness.png", height = 4.47, width = 7.2)

all_closeness

```


## Importance over the series

Finally we will visualise the trend in character importance accross the series

```{r importance evol, echo=TRUE, message=FALSE, warning=FALSE}

evol_all <- read_csv('Centrality Measures All/Importance.csv') %>%
  mutate(Book = c("Fellowship of the Ring", "The Two Towers", "Return of the King")) %>%
  select(Book, Frodo, Sam, Merry, Pippin, Aragorn, Legolas, Gimli, Gandalf) %>%
  melt(Value.name = "centrality") %>%
  rename(character = variable, centrality = value)
  

evol_plot <- ggplot(evol_all, 
                    aes(
                      x = factor(Book, levels = c("Fellowship of the Ring", "The Two Towers", "Return of the King")), 
                      y = centrality, 
                      group = character, 
                      color = character)
                    ) +
  
  scale_x_discrete(expand = c(0.05,0.05)) +
  
  #scale_y_continuous(limits = c(0, 0.72)) +
  
  geom_line(size = 2) +
  
  labs(x = NULL, y = "Centrality", title = "Character Importance by Book", subtitle = "Character importance by book, as measured by centrality", color = "Character") +
  
  theme_classic() +
  
  theme(
    
    panel.grid.major.y = element_line(color = "grey",
                                          size = 0.1,
                                          linetype = 1)
    )


ggsave("Images/all_evolve.png", height = 4.47, width = 7.2)

evol_plot

```
